# Introduction

```{r}
#| results: "asis"
#| echo: false
source("_common.R")
```

## Prerequisites

Following software need to be installed on your computer to work with the examples shown in this online book:

-   `R` Version 4.1 (or higher) <https://cloud.r-project.org/>. Check your R version by running following command in your console:
    ```{r r-version}
    R.Version()$version.string
    ```

-   `RStudio` Version 2022.02 or higher <https://www.rstudio.com/products/rstudio/download/#download>. Check your `RStudio` version by either looking clicking *About RStudio* under the menu *Help*, or by running following command in your console

    ```{r rstudio-version}
    #| eval: false
    
    rstudioapi::versionInfo()$version
    ```

Following software are only needed for specific chapters/examples:

-   `Git` <https://git-scm.com/downloads>
-   For Windows: `Rtools` \<https://cran.r-project.org/bin/windows/Rtools/

## Frequently used R packages

See also \@ref(Installing R packages)

Following R packages will be often used in the given examples and it is thus recommended 
to install them before starting with this book

-   `here`
-   `tidyverse`(installs `ggplot2, dplyr, tidyr, tibble, readr, forcats, stringr, purrr`)
-   `readxl`
-   `remotes`

Run this in your R command line to install these packages:

```{r}
#| eval: false

pkg_list <- c("here", "tidyverse", "readxl", "remotes")
install.packages(pkg_list)
```
