# PCA Plots

```{r}
#| label: setup
#| echo: false
#| message: false
source("_common.R")
library(knitr)

```

## Introduction

This tutorial is based on a nice blog post by Claus Wilke: <https://clauswilke.com/blog/2020/09/07/pca-tidyverse-style/>

Functions from the Tidyverse packages (`dplyr` , `tidyr`, `ggplot2`) and `broom` will be used in this tutorial. As example data a lipidomics dataset is imported that contains the lipid concentrations as columns and with one additional column indicating the experimental group

```{r}
#| label: load-lib
#| echo: true
#| message: false
library(here)
library(tidyverse)
library(broom)
library(cowplot)

```

```{r}
#| label: read-data
#| echo: true
#| message: false

d_long <- read_csv(here("data/Metabolites-1614644-supplementary.csv")) 
```

```{r}
#| label: tbl-data
#| echo: false
#| tbl-cap: "Test dataset in tidy format"
kable(head(d_long, n = 5),digits = 2)
```

## Transform and Scale the Data

```{r}
#| label: prcomp
#| echo: true
#| message: false

pca_res <- d_long %>% 
  select(where(is.numeric)) %>% # retain only numeric columns
  log2() |> 
  prcomp(scale = TRUE)

```

Compare this to plotting non-normalized concentrations

```{r}
#| label: pcaplot-1
#| echo: false
#| message: false

# Use broom to make the data ggplot ready and include annotations
pca_annot <- pca_res |> broom::augment(d_long)

ggplot(data = pca_annot, aes(.fittedPC1, .fittedPC2, color = Group)) + 
geom_point(size = 1.5) + 
stat_ellipse(level = 0.95) +
theme_light() +
theme(aspect.ratio=1)

```

```{r}
#| label: pcaplot-2
#| echo: false
#| message: false

# Use broom to make the data ggplot ready and include annotations
pca_annot <- pca_res |> broom::augment(d_long)

pca_12 <- ggplot(data = pca_annot, aes(.fittedPC1, .fittedPC2, color = Group)) + 
  geom_point(size = 1.5) + 
  stat_ellipse(level = 0.95)+
  theme_light() +
  theme(aspect.ratio=1) + theme(legend.position = "none")


pca_34 <- ggplot(data = pca_annot, aes(.fittedPC3, .fittedPC4, color = Group)) + 
  geom_point(size = 1.5) + 
  stat_ellipse(level = 0.95) +
  theme_light() +
theme(aspect.ratio=1) + theme(legend.position = "none")

cowplot::plot_grid(pca_12, pca_34, nrow = 1, rel_widths = c(1,1), labels = c("A","B"))

```