# Tables in R

```{r}
#| results: "asis"
#| echo: false
source("_common.R")
```

## Introduction

In this chapter we will look at how data is used and managed within R.

### Prerequisites

```{r}
#| label: setup
#| message: false
library(here)
library(tidyverse)
```

::: panel-tabset
## Base

```{r}
plot(cars)
```

## ggplot2

```{r}
ggplot(data = cars, aes(x = speed, y = dist)) + geom_point()
```
:::

## Important R data formats

+-------------+-----------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------+
| Type        | Description                                                                 | Example(s)                                                                                                    |
+=============+=============================================================================+===============================================================================================================+
| Vector      | Vector (series) of values. All values are of the same data type (see below) | `c(1,2,3,4)c("TQC", "BQC", "SPL")c(TRUE, FALSE)`                                                              |
+-------------+-----------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------+
| Matrix      | 2-dimensional set of values. All values are of the same data type           | `matrix(data = c(1:2),nrow = 2,ncol = 3)matrix(data = c("BQC", "SPL"),nrow = 2,ncol = 3)`                     |
+-------------+-----------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------+
| Data Frame\ | Table with columns that can have different data types                       | `tibble(   No = c(1,2,3),    Sample = c("SPL_1", "SPL_2", "SPL_3"))`                                          |
| Tibble      |                                                                             |                                                                                                               |
+-------------+-----------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------+
| List        | Series of objects                                                           | `list(   Studysite = c("NUH", "SGH"),   Cohort =      tibble(No = c(1,2),             Size = c(110,332)   ))` |
+-------------+-----------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------+

## Important R data types

| Type             | Short    | Description             | Example(s)           |
|------------------|----------|-------------------------|----------------------|
| Numeric (double) | `<dbl>`  | Floating point number   | `3.14159`            |
| Character        | `<char>` | Character (text) string | `"S1P d18:1", "BQC"` |
| Logical          | `<lgl>`  | True or False, 1 or 0   | `TRUE; FALSE`        |
| Integer          | `<int>`  | Number without digits   | `3L; 7011L; -13L`    |
| Factor           | `<fct>`  | Categorical data        | `BQC; TQC; SPL`      |

Moreover there are data and time types, i.e. date `<dt>`, time `<tm>` and datetime `<dttm>`

You can print tables in the console or in R notebook (.Rmd), which will also show the column types:
